---
title: "p8105_hw2_sl4269"
subtitle: "Data Wrangling 1"
author: "Steven Lawrence"
date: "September 28, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Problem 1

Loading libraries

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
```


## Importing the Data sets

Cleaning Mr.trash data

```{r data import, message=FALSE, warning=FALSE}
mr.trash <-
  readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", na = "", trim_ws = T, skip = 1, sheet = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster)%>%
  mutate(sports_balls = sports_balls %>% round(digits =0) %>% as.numeric())




```

Cleaning and merging precipitation data sets

```{r importing datasets, message=F}

##precipitation is in inches
precip17<- 
  readxl::read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", na= "", trim_ws = T, skip = 1, sheet = 6) %>% 
  janitor::clean_names() %>% 
  drop_na(total)%>%
  mutate(year = "2017", month = month.name[month])

precip18<- 
  readxl::read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", na= "", trim_ws = T, skip = 1, sheet = 5) %>% 
  janitor::clean_names() %>% 
  drop_na(total)%>%
  mutate(year = "2018", month = month.name[month])

precip18_and_17 <- rbind(precip17, precip18)

precip_by_year<- precip18_and_17 %>% 
  pivot_wider(
              names_from = year,
              values_from = total)
precip_by_year$month[13]<- "Total precipitation"

```

###Datadiscription

The `precip18_and_17` dataset contains precipitation data of 2017 and 2018. The dimentions of this dataset is `13x3` where the 13th row is total precipiation. For the available data the total precipitation of 2018 was `r pull(precip_by_year[13,3])` and the median number of sports balls in the dumpster in 2017 is `r mr.trash %>% pull(sports_balls) %>% median()`.

#Problem 2

```{r cleaning datasets, message=F, warning=F}

polls <- 
  read_csv("data/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, c('year', 'month', 'day') ) %>% 
  mutate(
    month = month.abb[as.numeric(month)] %>% tolower(), 
    president = prez_gop + prez_dem 
    ) %>% 
  select(year, month, president, everything(), -day, - prez_gop, -prez_dem) %>% 
  arrange(year, month)

snp <- 
  read_csv("data/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, c("month","day","year")) %>% 
  mutate( 
    month= month.abb[as.numeric(month)] %>% tolower()
  ) %>% 
  select(year, month, everything()) %>% 
  arrange(year,month)

unemployed <- 
  read_csv("data/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    year = as.character(year)
  ) %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment"
  ) %>% 
  arrange(year,month)

polls_snp<- merge(polls, snp, by = c("year","month")) 

polls_snp_umemployed <- left_join(polls_snp, unemployed, by = c("year", "month")) %>% 
  select(everything(), - day)

view(polls_snp_umemployed)

dim(polls_snp_umemployed
    )

```

The polls data set contains poll information from all the branches of government from the time of `1947 - 2015`. The dimenstion of this dataset is `822 by 9`. The snp data set contains information about the number of retracted peer reviewed papers from `1950 - 2015` with a dimension of `787 by 4`. Lastly, the unemployed dataset contains information of unemployment percentages from years `1948 - 2015` and dimentions of `816 to 3`. When the data sets combined, thet are filtered by the time period`1950 - 2015` and the dimensions are `786 by 11`.


# Problem 3

Loading in data set and cleaning it
```{r}

Baby_names <- 
  read_csv("./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  distinct() %>% 
  mutate(rank = as.factor(rank))
```

Table of Olivai's rank over time and ethnicity. 
```{r}

baby_olivia<- 
  Baby_names %>% 
  filter(childs_first_name =="Olivia") %>% 
  group_by(ethnicity, year_of_birth) 
  baby_olivia<- baby_olivia[order(pull(baby_olivia,rank)),] 
  baby_olivia %>% 
  knitr::kable(digits = 1)

```

Table of most popular male names.
```{r}
  
baby_popular_male_name <-
  Baby_names %>% 
  filter(gender == "MALE", rank == 1)

baby_popular_male_name <- 
  baby_popular_male_name[
    order(
      pull(
        baby_popular_male_name,-rank)
      ),
    ]

baby_popular_male_name %>% 
  knitr::kable(digits = 1)

```

Scatter plot of male, white non-hispanic, baby name observastions along wtih their rank. 

```{r}
baby_male_name_scatterplot <-
  Baby_names %>% 
  filter(gender == "MALE", ethnicity == "WHITE NON HISPANIC", year_of_birth == 2016) %>% 
  ggplot(aes(x = rank, y = count))+geom_point()+labs(title = "Male White Non Hispanic baby name count by Rank in 2016")

baby_male_name_scatterplot
```

